Time stretching of the GeV emission of GRBs: Fermi-LAT data vs geometrical model

### 2015-11-29 Paper/gammaRays.tex
Proceedings editing
Add more explanation to the section 5.1
Transition between general description of a fit, to the fit of GRB
090926A is explained.

	\begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{modelOverview}
        \caption{ Model Overview.  The red and blue cones represent
          the regions through which low and high energy charged
          particles propagate.  In the case shown the observer's
          off-axis angle is smaller than the opening angle of the low
          energy jet, so, due to the relativistic beaming effect, the
          most of the observable low energy photons will travel along
          the straight line from the central engine.  Also, the
          observer's off-axis angle is larger that the opening angle
          of the high energy jet, but the high energy radiation will
          still originate near the center of the jet as this is the
          only place where the high energy radiators are located.  The
          observation time of a photon is a sum of two terms. First, the
          time interval spent in plasma as a radiator which
          approximately equals to the distance from the central engine
          to the point of emission. Second, the time interval from
          emission to detection which equals to the distance from the point
          of emission to the observer's location.  For the given position
          of the observer the sum is larger for high energy
          photons.  Therefore, the high energy emission will be
          observed later throughout the burst duration and the
          high energy light curve will be stretched.  }
        \label{fig:modelOverview}
	\end{figure}

@@

First of all, we need to determine, which parameters to fit. At most,
the observables of a particular burst depend on 10 parameters: 8 of
the burst $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k,
\alpha\right)$, and 2 of the observer's position related to the burst
$\left(z, \chi\right)$. However, some of these parameters only change
observables trivially, and, also, there are transformations of
variables, which do not change observables at all (see equation
\ref{eq:p}):
\begin{itemize}
	\item{The number of photons $p$ depends linearly on $\eta_0$,
          so $\eta_0$ doesn't affect other observables like duration,
          or stretching factor. One can easily fit $\eta_0$ to match
          the observed photon count.}
	\item{If we make a simultaneous transformation of $r_0
          \rightarrow \lambda r_0$ and $\eta_0 \rightarrow
          \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter
          of transformation, only duration of the burst will
          change. Other observables, like the total number of photons,
          or stretching factor, will not be affected.}
	\item{Finally, if we make a following transformation:
          $\omega_0 \rightarrow \lambda \omega_0$, $\theta_0
          \rightarrow \lambda^k \theta_0$, $\eta_0 \rightarrow
          \lambda^\alpha \eta_0$, where $\lambda$ is another
          transformation parameter, no observables will change
          whatsoever.}
\end{itemize}
Given these transformations, one can reduce the number of parameters
for fit to 7: $\left(\gamma, n, \theta_0, k, \alpha, z, \chi\right)$.

@@

The best fit parameters are found by minimizing the cost function
$C\left(n, \theta_0, k, \alpha, \chi\right)$ which should
satisfy the following objectives:
\begin{itemize}
	\item { Total energy of the burst is finite (that is
          $k+\frac{\alpha}{2}+1 < 0$) and smaller than $\unit[6 \times
            10^{53}]{GeV} \approx \unit[10^{51}]{erg}$
          \cite{Gehrels:2013xd}.  }
	\item { The stretching factor of the burst should be
          compatible with the value for the GRB 090926A.  }
	\item { The ratio of photons with high energies and low
          energies should be compatible with the value for the GRB
          090926A, $\frac{p_{\infty}
            \left(z,\chi;\unit[1]{GeV},\infty\right)}{p_{\infty}
            \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)} =
          0.057$.  }
	\item { The burst should not be too faint compared to
          other bursts from the sample. In other words, the total
          number of observed photons should approximately equal to the
          median among the sample.  }
	\item { We assume that all bursts have the same burst
          parameters $\left(\gamma, \eta_0, r_0, n, \omega_0,
          \theta_0, k, \alpha\right)$ and require that bursts with
          small stretching factors like GRB 090902B should be possible
          to observe. So we require that bursts with the stretching
          factor of GRB 090902B or lower appear in random samples by
          varying $z$ and $\chi$ (sampling is done the same way as in
          Section \ref{sec:distribution}). Note that this requirement
          unlike others applies to an ensemble of simulated bursts. }
\end{itemize}

>>>>

	\begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{modelOverview.pdf}
        \caption{ Model Overview.  The red and blue cones represent
          the regions through which low and high energy charged
          particles propagate.  In the case shown the observer's
          off-axis angle is smaller than the opening angle of the low
          energy jet, so, due to the relativistic beaming effect, the
          most of the observable low energy photons will travel along
          the straight line from the central engine.  Also, the
          observer's off-axis angle is larger that the opening angle
          of the high energy jet, but the high energy radiation will
          still originate near the center of the jet as this is the
          only place where the high energy radiators are located.  The
          observation time of a photon is a sum of two terms. First, the
          time interval spent in plasma as a radiator which
          approximately equals to the distance from the central engine
          to the point of emission. Second, the time interval from
          emission to detection which equals to the distance from the point
          of emission to the observer's location.  For the given position
          of the observer the sum is larger for high energy
          photons.  Therefore, the high energy emission will be
          observed later throughout the burst duration and the
          high energy light curve will be stretched.  }
        \label{fig:modelOverview}
	\end{figure}

@@

First of all, we need to determine, which parameters to fit. At most,
the observables of a particular burst depend on 10 parameters: 8 of
the burst $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k,
\alpha\right)$, and 2 of the observer's position related to the burst
$\left(z, \chi\right)$. However, some of these parameters only change
observables trivially, and, also, there are transformations of
variables, which do not change observables at all (see equation
\ref{eq:p}):
\begin{itemize}
	\item{The number of photons $p$ depends linearly on $\eta_0$,
          so $\eta_0$ doesn't affect other observables like duration,
          or stretching factor. One can easily fit $\eta_0$ to match
          the observed photon count.}
	\item{If we make a simultaneous transformation of $r_0
          \rightarrow \lambda_1 r_0$ and $\eta_0 \rightarrow
          \frac{1}{\lambda_1^3}\eta_0$, where $\lambda_1$ is a parameter
          of transformation, only duration of the burst will
          change. Other observables, like the total number of photons,
          or stretching factor, will not be affected.}
	\item{Finally, if we make a following transformation:
          $\omega_0 \rightarrow \lambda_2 \omega_0$, $\theta_0
          \rightarrow \lambda_2^k \theta_0$, $\eta_0 \rightarrow
          \lambda_2^\alpha \eta_0$, where $\lambda_2$ is another
          transformation parameter, no observables will change
          whatsoever.}
\end{itemize}
Given these transformations, one can reduce the number of parameters
for fit to 7: $\left(\gamma, n, \theta_0, k, \alpha, z, \chi\right)$.

@@

We are going to fit parameters specifically for GRB 090926A for its stretching factor is the most significant. While we assume burst related parameters $\left(n, \theta_0, k, \alpha\right)$ to be the same for all bursts, $\chi$ here considered to be the off-axis angle of GRB 090926A. At the same time, we need to make sure that burst related universal parameters $\left(n, \theta_0, k, \alpha\right)$ are consistent with observation of other bursts such as GRB 090902B.

In particular, the best fit parameters are found by minimizing the cost function
$C\left(n, \theta_0, k, \alpha, \chi\right)$ which
satisfies the following objectives:
\begin{itemize}
	\item { Total energy of the burst is finite (that is
          $k+\frac{\alpha}{2}+1 < 0$) and smaller than $\unit[6 \times
            10^{53}]{GeV} \approx \unit[10^{51}]{erg}$
          \cite{Gehrels:2013xd}.  }
	\item { The stretching factor of the burst should be
          compatible with the value for the GRB 090926A.  }
	\item { The ratio of photons with high energies and low
          energies should be compatible with the value for the GRB
          090926A, $\frac{p_{\infty}
            \left(z,\chi;\unit[1]{GeV},\infty\right)}{p_{\infty}
            \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)} =
          0.057$.  }
	\item { The burst should not be too faint compared to
          other bursts from the sample. In other words, the total
          number of observed photons should approximately equal to the
          median among the sample.  }
	\item { We assume that all bursts have the same burst
          parameters $\left(\gamma, \eta_0, r_0, n, \omega_0,
          \theta_0, k, \alpha\right)$ and require that bursts with
          small stretching factors like GRB 090902B should be possible
          to observe. So we require that bursts with the stretching
          factor of GRB 090902B or lower appear in random samples by
          varying $z$ and $\chi$ (sampling is done the same way as in
          Section \ref{sec:distribution}). Note that this requirement
          unlike others applies to an ensemble of simulated bursts. }
\end{itemize}
